Symbolic Dynamic Programming for Continuous State and Observation POMDPs

### 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex
updated draft4

\section{Introduction and Related Work}
% Write intro here

@@

For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
% For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
{\footnotesize
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=\open) &: t_s - 5 \\ 
(a \neq \open) &: t_s + 7 \\
\end{cases}
\right]\nonumber
\hspace{5mm}
R(t_s,a) = 
\begin{cases}
 (a=\open) &: -1 \\
(a \neq \open)\wedge (t_s>T) &: -1000 \\
(a \neq \open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}\nonumber
\end{align}
}
%contribution? 
This example will try to maximize its reward using observations on the temperature. The observations received before closing the plant to a safe temperature are therefore critical.  For the observation model we begin with a discrete model for the observation $z$ with values $high, low$ such as the following for the $\open$ action:  
{\footnotesize
\begin{align}
P(z=high|t_s',\open) = 
\begin{cases}
  t_s <= 15 &: 0.9 \\
 \neg (t_s <= 15)&: 0.1 \\
\end{cases}\nonumber
\hspace{5mm}
P(z=low|t_s',\open) = 
\begin{cases}
 t_s <= 15 &: 0.1 \\
 \neg (t_s <= 15) &: 0.9 \\
\end{cases}\nonumber
\end{align}
}
%\\
%P(t_{o_1}|t_s',\close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.1 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
%\end{cases}\nonumber
%,
%P(t_{o_2}|t_s',\close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.9 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
%\end{cases} \nonumber
%\end{align}
After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains. %For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
{\footnotesize
\begin{align}
P(t_o|t_s',\open) = 
\begin{cases}
 (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
 \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
\end{cases}\nonumber
\end{align}
}
Next we define the general solution to solving DC-POMDPs.

@@

At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$. \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
}
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
%The decision at each time step is to pick an action based on past actions and observations. 
The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1. %From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} 
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
              $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

       \If{$V^h = V^{h-1}$}
           {break $\,$ \emph{// Terminate if early convergence}\;}
   }
     \Return{$(V^h,\pi^{*,h})$} \;
}
\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.

Point-based dynamic programming backups compute the optimal value function given a belief state $b$. The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups: 
{\footnotesize
\begin{align*}
\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
\end{align*}
}
where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$.

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
{\footnotesize
\begin{align}
g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
}
where $r_a = r(x_s,d_s,a)$.  
%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
Lines 7--15 of the \textsc{PBVI} algorithm show the above in a general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}. A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.

@@

\section{Symbolic Dynamic Programming} 
Dynamic programming provides exact solutions to POMDP problems for small domains. As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons. Avoiding this enumeration often requires approximating the solutions. We claim an exact solution using symbolic dynamic programming. The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors. We define the requirements to implement the value iteration algorithm for DC-POMDP.

\subsection{Case Representation and Extended ADDs}
% operations, max, restrict, substitute
%overview + example plant
The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
{\footnotesize 
\begin{align}
f = 
\begin{cases}
  \phi_1: & f_1 \\ 
 \vdots&\vdots\\ 
  \phi_k: & f_k \\ 
\end{cases} \nonumber
\end{align}
}
Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.  
The $f_i$ are function definitions either linear or quadratic in the continuous parameters.

@@

@@

@@

The next three operations are required in defining the value iteration algorithm. A \emph{symbolic case maximization} is defined as below:

@@

%xadd representation
The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
case statements and the required operations.  Here we use continuous SDP for both states and observations in DC-POMDPs. The $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs. For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD. Figure 1 (right) is an example of an XADD representation.

@@

Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations. %To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons. 
Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2 of this paper. Also a belief is given as the input and often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in \mathbb{B}$.

For each iteration, lines 11--12 product the following equation: 
{\footnotesize
\begin{align}
g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
\end{align}
}
where the $\alpha$-vector set are case functions primed by substitution in \textsc{Prime} 
and \texttt{Backup} performs regression on the $\alpha$-vectors. Boolean regression is performing a symbolic restriction while a continuous regression is a symbolic substitution as defined previously. 
Line 13 takes the cross-sum of the two sets which takes every possible combination of observations according to the $\alpha$-vectors. 
All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
The same operation is performed for all actions and then in line 15 the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon. The action from which this $\alpha$-vector was chosen from is the maximum policy. 
This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.

@@

Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$. It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
This operation of line 3 in \textsc{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model. For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section. 
Assuming that after $h=1$ the set of $\alpha$-vectors are defined as:
{\footnotesize
\begin{align}
\alpha_1^1(s) &= 
\begin{cases}
 (t_s>200) &: -1000 \\
\neg(t_s>200) &: 100 \\
\end{cases}
\hspace{10mm} 
\alpha_2^1(s) = \top: -1 \nonumber
\end{align}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}

>>>>

\section{Introduction} %and Related Work}
% Write intro here

@@

For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
% For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
{\footnotesize
\vspace{-1mm}
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=\open) &: t_s - 5 \\ 
(a \neq \open) &: t_s + 7 \\
\end{cases}
\right]
\hspace{2mm}
R(t_s,a) = 
\begin{cases}
 (a=\open) &: -1 \\
(a \neq \open)\wedge (t_s>T) &: -1000 \\
(a \neq \open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}
\end{align}
\vspace{-4mm}
}

%contribution? 
This example will try to maximize its reward using observations on the temperature. The observations received before closing the plant to a safe temperature are therefore critical.  For the observation model we begin with a discrete model for the observation $z$ with values $high, low$ such as the following for the $\open$ action:  
{\footnotesize
\vspace{-1mm}
\begin{align}
P(z=high|t_s',\open) = 
\begin{cases}
  t_s <= 15 &: 0.9 \\
 \neg (t_s <= 15)&: 0.1 \\
\end{cases}
\hspace{5mm}
P(z=low|t_s',\open) = 
\begin{cases}
 t_s <= 15 &: 0.1 \\
 \neg (t_s <= 15) &: 0.9 \\
\end{cases}
\end{align}
\vspace{-3mm}
}
%\\
%P(t_{o_1}|t_s',\close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.1 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
%\end{cases}\nonumber
%,
%P(t_{o_2}|t_s',\close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.9 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
%\end{cases} \nonumber
%\end{align}

After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains. %For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
{\footnotesize
\begin{align}
P(t_o|t_s',\open) = 
\begin{cases}
 (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
 \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
\end{cases}
\end{align}
}
Next we define the general solution to solving DC-POMDPs.

@@

At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)\mathcal{T}(xd'_s,xd_s,a)\mathcal{Z}(xd'_s,xd_o,a)$. \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
}
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
%The decision at each time step is to pick an action based on past actions and observations. 
The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$b_i \in B$}
       {
       	\ForEach {$a \in A$}
      	 {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
       	 }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
      }
             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
             % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
      \ForEach {$b_i \in B$}
      {
     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
     	$\Gamma_{PBVI}^h :=\ \Gamma^h \cup \alpha_{b_i}^h$\;
      }

       \If{$V^h = V^{h-1}$}
           {break $\,$ \emph{// Terminate if early convergence}\;}
   }
     \Return{$\Gamma_{PBVI}$} \;
}
\caption{\footnotesize \texttt{PBVI}(DC-POMDP, $H$,$ContObs$, $B=\left\{b_i \right\}$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.

\textbf{Point-based dynamic programming backups} compute the optimal value function given a belief state $b$. The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups: 
{\footnotesize
\begin{align*}
\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
\end{align*}
}
where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$.

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
{\footnotesize
\begin{align}
g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
}
where $r_a = r(x_s,d_s,a)$.  
%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
Lines 7--15 of the \textsc{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}. A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.

@@

\section{Symbolic Dynamic Programming} 
Dynamic programming provides exact solutions to POMDP problems for small domains. As the domain grows, fully enumerating the state, action and observation space leads to intractable solutions even for low horizons. Avoiding this enumeration often requires approximating the solution. We claim an exact solution using symbolic dynamic programming. The SDP solution for fully observable domains has been presented in \cite{sanner_uai11} %using piecewise constant functions
Our symbolic algorithm deals with continuous observations and states by generating the relevant observation partition at every time step and performing the backup operation using $\alpha$-vectors. We define the requirements to implement the value iteration algorithm for DC-POMDP.

\subsection{Case Representation and Extended ADDs}
% operations, max, restrict, substitute
%overview + example plant
The \textsc{Power Plant} example represented all functions using the case structure which can be generally defined as:
{\footnotesize 
\vspace{-3mm}
\begin{align}
f = 
\begin{cases}
  \phi_1: & f_1 \\ 
 \vdots&\vdots\\ 
  \phi_k: & f_k \\ 
\end{cases} \nonumber
\end{align}
}
Where $\phi_i$ are disjoint logical formulae defined over the state $(\vec{d},\vec{x})$ with logical ($\land,\lor,\neg$) combinations of boolean variables and inequalities ($\geq,>,\leq,<$) over continuous variables.  
The $f_i$ are function definitions either linear or quadratic in the continuous parameters.

@@

\vspace{-3mm}

@@

\vspace{-2mm}

@@

A \emph{symbolic case maximization} is defined as below:

@@

%xadd representation
The data structure of \emph{extended ADD} (XADD)~\cite{sanner_uai11} is used to support
case statements and the required operations.  Here we use continuous SDP for both states and observations in DC-POMDPs. The $\alpha$-vectors partition into states of equal value, both of which can be represented using XADDs. For increasing efficiency linear constraint feasibility checkers such as LP solvers allow pruning unreachable paths in the XADD. Figure 1,right is an example of an XADD representation.

@@

Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations. %To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons. 
Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2. Also a set of beliefs are given as the input as often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in B$.
For each iteration, lines 11--12 product the following equation: 
{\footnotesize
\vspace{-2mm}
\begin{align}
g^h_{a,o,j} =  \int_{x_{s'}} \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
\end{align}
}
where the first probability is from equation (2) and the next probability is from equation (1) of the example. The $\alpha$-vectors are case functions primed using substitution in \textsc{Prime}. \texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations. 
Lines 14--16 cover closed form operations on the $\Gamma$-set such as cross-sum and union. Finally lines 18--20 find the maximum $\alpha$-vector within the $\Gamma^h$ set and add this to the final $\Gamma_{PBVI}^h$ if it is not already in the set. 
%Line 13 takes the cross-sum of the two sets, followed by adding the reward to the discounted result. For all actions the  $\Gamma$-set is added to the $\Gamma^h$ of the current horizon. The result is pruned as explained earlier.%which takes every possible combination of observations according to the $\alpha$-vectors. 
%All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
%The same operation is performed for all actions and then in 
%Line 15 computes the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
%The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon. The action from which this $\alpha$-vector was chosen from is the maximum policy. 
%This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.

@@

Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$. It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model. For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section. 
Assuming that after $h=1$ the set of $\alpha$-vectors are defined as:
{\footnotesize
\vspace{-3mm}
\begin{align}
\alpha_1^1(t_s) &= 
\begin{cases}
 (t_s>200) &: -1000 \\
\neg(t_s>200) &: 100 \\
\end{cases}
\hspace{10mm} 
\alpha_2^1(t_s) = \top: -1 \nonumber
\end{align}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}
